Characterizing the response of subarctic European vegetation to key environmental variables for land surface modeling of ozone pollution and risk assessment

### 2021-10-15 VegOzone.tex
Checked section 2 with grammerly

We performed ozone concentration measurements in the subarctic, during the growing seasons of 2018 and 2019. Our chosen location was Svanhovd ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m\,a.s.l.}$) for several reasons. There were ozone measurements there in the past, there is an agrometeorological station measuring many of the climate variables we needed, e.g. $2\,\unit{m}$ temperature, precipitation, global irradiance and soil temperature at different depths. We collected atmospheric monitoring and ozone monitoring data taken at Svanhovd over the past $35\,\unit{years}$. There is also ongoing national surveillance of air pollution (\chem{NO_x}, \chem{SO_x}, PM, radiological species). Most of these measurements are performed by the Norwegian Institute for Air Research (NILU) and they also performed our ozone measurements.\\

All measured essential climate variables for the 2018/19 growing season were obtained through \href{luftkvalitet.no}{luftkvalitet.no} operated by \citet{NILU_AIRQ} (last accessed May 2020). Our ozone monitoring data have been added to luftkvalitet.no and the EBAS database operated by \citet{NILU_EBAS} and are thus openly accessible. Long-term ozone observation data (1986--1996) were obtained from EBAS. Note that data from before the 1990s do not follow the high-quality standard procedures implemented nowadays and have to be treated with care \citep{NILU2003}. Especially, ozone monitors did not undergo regular re-calibration. This probably led to drifts in the observed data and may impose a false trend which shall not be of concern for us as we are interested in the seasonality.
The long-term data sets of agrometeorological variables including temperature and precipitation are available from September 1992 to the present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.

@@

\begin{figure}[t]
  \centering
  (a)\\
  \includegraphics[width=8.3cm]{svanhovd_researchstation}\\
  (b)\\
  \includegraphics[width=8.3cm]{IMG_7739.JPG}
  %\includegraphics[width=8.3cm]{fig01}
  \caption{NIBIO Environment Centre Svanhovd close by the settlement of Svanvik, Norway. (a) Atmospheric monitoring site and ozone garden have been marked. Aerial photography \copyright Norges Kartverk; (b) Clover in the Svanhovd ozone garden. The plants had to be secured against herbivores with a wire-mesh fence. The plants shown are approximately $6-12\,\unit{cm}$ high.}
  \label{fig:svanhovd_research_station}
\end{figure}

@@

Hourly averaged $2\,\unit{m}$ temperatures above $20\,\unit{^\circ C}$ occurred more regularly in July 2018 than in 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}b). In 2018, spring temperature regularly rose above freezing only in May, while in 2019 this occurred already early in March/April.
More rain events with accumulated daily precipitation ($\sum_d \mathrm{Precip}$) above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c).
Qualitatively, global irradiance ($Q_0$) displayed in Fig.~\ref{fig:data_svanvik_2018_2019}d was higher in May and July 2018 compared to 2019, while June 2019 showed higher irradiance than in 2018. A high global irradiance in most cases is the result of a low cloud fraction. In both years, the maximum recorded $Q_0$ was $750\,\unit{W\,m^{-2}}$ and reached in June.

@@

Svanhovd is located within the subarctic climate zone. The 1992--2012 climatologies of temperature and precipitation are shown in a combined climate diagram (Fig.~\ref{fig:climatediagram}). Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with a standard error of the mean (SE) as error band. We chose SE in this case due to the large interannual variability in temperature reflected in the standard deviation. Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation as error bars. Months with average temperatures below freezing are denoted with a star. As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppen's climate classification of a regular subarctic climate (Dfc) \citep[e.g.][]{SD:Beck2018}. The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February. The coldest measured temperature was $-45.2\,\unit{^\circ C}$ in January 27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred in July 16 the same year.

@@

In Fig.~\ref{fig:anomalies_svanvik}a, temperature anomalies for Svanvik are shown as a percentage of days warmer or colder than $\pm 1\,\sigma$ from the climatology for each month. Negative deviations from climatology are displayed as a negative percentage. The annual positive/negative deviations are indicated in the respective corners (right upper/right lower). The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$). We find, that the summer of 2018 was significantly warmer than average. Especially in May and July more than $40\,\unit{\%}$ of the days were significantly warmer. Significantly more warmer days continued to occur in all months (August--November) except for October. March 2018 had many unusually cold days. The summer of 2019 was fairly average, however, significantly more colder days occurred in July, while August had both significantly colder and warmer days. April 2019 had significantly warmer than average days.

@@

%Ozone
We use the bias-corrected and cross-calibrated ozone climatology \citep{ACPD:Falk2021} and assess the monthly significance of the ozone concentration anomalies in 2018/19. The resulting percentage of days above/below $\pm 1\,\sigma$ is displayed for each month in Fig.~\ref{fig:anomalies_svanvik}d. Note that missing bars for months in winter and spring are both due to the temporal extend of our data and anomalies not meeting the $1\sigma$ criteria. The star indicates the reconstructed data in July. In April/May 2018, ozone was significantly elevated, while no significant enhancement is found in 2019. In both years, ozone concentrations were significantly lower than average in November. These results suggest that stagnant meteorological conditions in April/May in combination with a late start of the growing season contributed more to elevated ozone than the extensive forest fires in Sweden that did not significantly enhance \chem{[O_3]} in 2018.

>>>>

We performed ozone concentration measurements in the subarctic, during the growing seasons of 2018 and 2019. Our chosen location was Svanhovd ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m\,a.s.l.}$) for several reasons: (1) Ozone had been monitored there in the past, (2) an agrometeorological station is measuring, e.g. $2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths, and (3) a national surveillance of air pollution (\chem{NO_x}, \chem{SO_x}, PM, radiological species) is ongoing. Most of these measurements are performed by the Norwegian Institute for Air Research (NILU), who also conducted ozone measurements on our behalf. In addition, we collected atmospheric monitoring and ozone monitoring data taken at Svanhovd over the past $35\,\unit{years}$.\\

All measured essential climate variables for the 2018/19 growing season were obtained through \href{luftkvalitet.no}{luftkvalitet.no} operated by \citet{NILU_AIRQ} (last accessed May 2020). Our ozone monitoring data have been added to luftkvalitet.no and the EBAS database operated by \citet{NILU_EBAS} and are thus openly accessible. Long-term ozone observation data (1986--1996) were obtained from EBAS. Note that data from before the 1990s do not follow the high-quality standard procedures implemented nowadays and have to be treated with care \citep{NILU2003}. Especially, ozone monitors did not undergo regular re-calibration. That probably led to drifts in the observed data and may impose a false trend which shall not be of concern for us as we are interested in the seasonality.
The long-term data sets of agrometeorological variables including temperature and precipitation are available from September 1992 to the present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.

@@

\begin{figure}[t]
  \centering
  (a)\\
  \includegraphics[width=8.3cm]{svanhovd_researchstation}\\
  (b)\\
  \includegraphics[width=8.3cm]{IMG_7739.JPG}
  %\includegraphics[width=8.3cm]{fig01}
  \caption{NIBIO Environment Centre Svanhovd close by the settlement of Svanvik, Norway. (a) Atmospheric monitoring site and ozone garden have been marked. Aerial photography \copyright Norges Kartverk; (b) Clovers in the Svanhovd ozone garden. The plants had to be secured against herbivores with a wire-mesh fence. The plants shown are approximately $6-12\,\unit{cm}$ high.}
  \label{fig:svanhovd_research_station}
\end{figure}

@@

Hourly averaged $2\,\unit{m}$ temperatures above $20\,\unit{^\circ C}$ occurred more regularly in July 2018 than in 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}b). In 2018, spring temperature regularly rose above freezing only in May, while in 2019 this occurred already early in March/April.
More rain events with accumulated daily precipitation ($\sum_d \mathrm{Precip}$) above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c).
Qualitatively, global irradiance ($Q_0$) displayed in Fig.~\ref{fig:data_svanvik_2018_2019}d was higher in May and July 2018 compared to 2019, while June 2019 showed higher irradiance than in 2018. A high global irradiance, in most cases, is the result of a low cloud fraction. In both years, the maximum recorded $Q_0$ was $750\,\unit{W\,m^{-2}}$ and reached in June.

@@

Svanhovd is located within the subarctic climate zone. Climatologies of temperature and precipitation derived for 1992--2012 are shown in a combined climate diagram (Fig.~\ref{fig:climatediagram}). Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with a standard error of the mean (SE) as error band. We chose SE in this case due to the large interannual variability in temperature reflected in the standard deviation. Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation as error bars. Months with average temperatures below freezing are denoted with a star. As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppen's climate classification of a regular subarctic climate (Dfc) \citep[e.g.][]{SD:Beck2018}. The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February. The coldest measured temperature was $-45.2\,\unit{^\circ C}$ in January 27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred on July 16 the same year.

@@

In Fig.~\ref{fig:anomalies_svanvik}a, temperature anomalies for Svanvik are shown as a percentage of days warmer or colder than $\pm 1\,\sigma$ from the climatology for each month. Negative deviations from climatology are displayed as a negative percentage. The annual positive/negative deviations are indicated in the respective corners (right upper/right lower). The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$). We find, that the summer of 2018 was significantly warmer than average. Especially in May and July more than $40\,\unit{\%}$ of the days were significantly warmer. A significant number of warmer days continued to occur in all months (August--November) except for October. March 2018 had many unusually cold days. The summer of 2019 was fairly average, however, more significantly colder days occurred in July, while August had both significantly colder and warmer days. April 2019 had significantly warmer than average days.

@@

%Ozone
We use the bias-corrected and cross-calibrated ozone climatology \citep{ACPD:Falk2021} and assess the monthly significance of the ozone concentration anomalies in 2018/19. The resulting percentage of days above/below $\pm 1\,\sigma$ is displayed for each month in Fig.~\ref{fig:anomalies_svanvik}d. Note that missing bars for months in winter and spring are both due to the temporal extent of our data and anomalies not meeting the $1\sigma$ criteria. The star indicates the reconstructed data in July. In April/May 2018, ozone was significantly elevated, while no significant enhancement is found in 2019. In both years, ozone concentrations were significantly lower than average in November. These results suggest that stagnant meteorological conditions in April/May in combination with a late start of the growing season contributed more to elevated ozone than the extensive forest fires in Sweden that did not significantly enhance \chem{[O_3]} in 2018.
